---
title: "DRiDO Figure 3"
author: "Johanna Fleischman"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide 
---

# Setup 

## Install Packages

```{r, message=F, warning=F}
suppressMessages(library(remotes))
suppressMessages(library(patchwork))
suppressMessages(library(flexclust))
suppressMessages(library(ggrepel))
suppressMessages(library(RColorBrewer))
suppressMessages(library(VennDiagram))
suppressMessages(library(ggh4x))
suppressMessages(library(ggpmisc))
suppressMessages(library(tidyverse))
suppressMessages(library(mclust))
suppressMessages(library(clusterProfiler))
suppressMessages(library(org.Mm.eg.db))
suppressMessages(library(rrvgo))
suppressMessages(library(GGally))
suppressMessages(library(grid))

knitr::opts_chunk$set(echo = TRUE, eval = TRUE, out.height = "100%", out.width = "100%", warning = F, message = F)
```

## Assign filepaths

```{r}
# set
# local_filepath <- "~/drido-multiomic-paper"

source(file.path(local_filepath, "R/statistics_functions.R"))
source(file.path(local_filepath, "R/figure_functions.R"))
source(file.path(local_filepath, "R/metab_enrichment_IDSLGOA.R"))

# Filepaths to gene ontology names and INCHI keys for metabolomics enrichment
go_names_filepath <- file.path(local_filepath, "data/go_names.RData")
go_inchi_key_filepath <- file.path(local_filepath, "data/all_go_child_cpd_list_met_ik14.RData")

lipidomics_data <- "inst/extdata/20250128-Normalized-Lipidomics-Data.Rds"
metabolomics_data <- "inst/extdata/20250128-Normalized-Metabolomics-Data.Rds"
proteomics_data <- "inst/extdata/20250129-Normalized-Proteomics-Data.Rds"
phenotype_data <- "inst/extdata/DOCR_Phenotype_Data.csv"

# PLL Modals on z-scored / scaled data
pll_lm_res_file <- "inst/extdata/20260220-Linear-Scaled-MM-PLL-Aging-Results.Rds"
pll_lm_death_res_file <- "inst/extdata/20260220-Linear-Scaled-MM-PLL-Dying-Results.Rds"
pll_lm_res_phys_file <- "inst/extdata/20250528-Linear-MM-PLL-Aging-Phenotype-Results.Rds"
pll_lm_death_res_phys_file <- "inst/extdata/20250528-Linear-MM-PLL-Dying-Phenotype-Results.Rds"

# PLL Models on centered, NOT scaled data. Only appropriate for use with panel D
# pll_lm_res_file <- "inst/extdata/20250424-Linear-MM-PLL-Aging-Results.Rds"
# pll_lm_death_res_file <- "inst/extdata/20250424-Linear-MM-PLL-Dying-Results.Rds"
# pll_lm_res_phys_file <- "inst/extdata/20250424-Linear-MM-PLL-Aging-Phenotype-Results.Rds"
# pll_lm_death_res_phys_file <- "inst/extdata/20250424-Linear-MM-PLL-Dying-Phenotype-Results.Rds"

pll_gam_summary_file <- "inst/extdata/20251008-PLL-GAM-Summary-Data.Rds"
pll_gam_fit_file <- "inst/extdata/20251008-PLL-GAM-Fit-Data.Rds"
name_conv_file <- "inst/supp_tables/Table_S1_CompoundAnnotations.csv"
physio_key_file <- "data/physio-key.csv"
inchii_key_file <- "data/inchii-key.csv"
```

## Load and clean data

```{r}
# Get keys
name_conversion_use <- read.csv(file.path(local_filepath, name_conv_file)) %>%
  dplyr::select(-modality, -Broad_Group)
physio_key <- read.csv(file.path(local_filepath, physio_key_file), header = TRUE)
inchii_key <- read.csv(file.path(local_filepath, inchii_key_file))

# Make final DO-CR raw data frame
# Retains pre-diet intervention data
data_use <- docr_make_final_data(
  metabolomics_data_filepath = file.path(local_filepath, metabolomics_data),
  proteomics_data_filepath = file.path(local_filepath, proteomics_data),
  lipidomics_data_filepath = file.path(local_filepath, lipidomics_data),
  phenotype_data_filepath = file.path(local_filepath, phenotype_data),
  name_conversion_key = name_conversion_use
)

# Import PLL Models
pll_lm_pheno_res <- readRDS(file.path(local_filepath, pll_lm_res_phys_file))
pll_lm_death_pheno_res <- readRDS(file.path(local_filepath, pll_lm_death_res_phys_file))

pll_lm_aging <- readRDS(file.path(local_filepath, pll_lm_res_file)) %>%
  dplyr::select(-method, compoundName) %>%
  dplyr::bind_rows(pll_lm_pheno_res)

pll_lm_dying <- readRDS(file.path(local_filepath, pll_lm_death_res_file)) %>%
  dplyr::select(-method, compoundName) %>%
  dplyr::bind_rows(pll_lm_death_pheno_res)

pll_gam_summaries <- readRDS(file.path(local_filepath, pll_gam_summary_file))
pll_gam_fits <- readRDS(file.path(local_filepath, pll_gam_fit_file))

pll_gam_cluster_identities <- pll_gam_summaries %>%
  dplyr::distinct(trait, cluster)

# Clean PLL Linear Models
pll_lm_long <- dplyr::bind_rows(
  pll_lm_aging %>%
    dplyr::mutate(
      model_term = ifelse(model_term == "PLL", "PLL_Aging", as.character(model_term)),
      model = gsub("PLL", "PLL_Aging", model),
      phase = "Aging"
    ),
  pll_lm_dying %>%
    dplyr::mutate(
      model_term = ifelse(model_term == "PLL", "PLL_Dying", as.character(model_term)),
      model = gsub("PLL", "PLL_Dying", model),
      phase = "Dying"
    )
) %>%
  dplyr::left_join(., name_conversion_use, by = "feature_id") %>%
  dplyr::left_join(., pll_gam_cluster_identities, by = c("feature_id" = "trait")) %>%
  dplyr::rows_update(
    physio_key %>% dplyr::select(feature_id, name_use),
    by = c("feature_id")
  ) %>%
  dplyr::filter(!is.na(name_use)) %>% # Remove bad annotations before FDR
  dplyr::mutate(
    model = make_names_pretty(model),
    model_term = make_names_pretty(model_term)
  ) %>%
  dplyr::select(-padj) %>%
  fdr_multi(
    pval_var = "p_value_coef",
    nest_vars = c("phase", "model_term", "modality")
  )


rm(list = c("pll_lm_pheno_res", "pll_lm_death_pheno_res"))
```

### Save Processed Data

```{r}
processed_data_export <- data_use %>%
  dplyr::rows_update(
    physio_key %>% dplyr::select(trait_id = feature_id, name_use),
    by = "trait_id"
  )

# saveRDS(processed_data_export, file.path(local_filepath, "supplemental_tables_final", "all_processed_data.RDS"))
```

### Tables: S3-4

```{r}
# PLL non-linear trajectory summaries
table_s3 <- pll_gam_summaries %>%
  as.data.frame() %>%
  dplyr::rows_update(
    physio_key %>% dplyr::select(trait = feature_id, name_use),
    by = c("trait")
  ) %>%
  dplyr::mutate(cluster = ifelse(is.na(cluster),
    "None assigned",
    cluster
  )) %>%
  dplyr::mutate(Platform = stringr::str_to_title(modality)) %>%
  dplyr::rows_update(
    physio_key %>% dplyr::select(trait = feature_id, name_use),
    by = c("trait")
  ) %>%
  dplyr::select(Platform,
    Compound = name_use, Cluster = cluster,
    pvalue, Smooth_Term = smooth_term,
    Pvalue = pvalue, FDR = padj,
    Compound_PMC = extrema_trait,
    Cluster_PMC = extrema_mean,
    Cluster_PMC_SE = extrema_se,
    Deviation_Exp = deviation.exp,
    Adj_R2 = adj.rsquared, EDF = edf
  ) %>%
  dplyr::arrange(Cluster, FDR)

# write.csv(table_s3,
#           file.path(local_filepath, "supplemental_tables_final", "Table_S3_PLL_GAM_Summary.csv"),
#           row.names = FALSE)


# Build table s4 for both scaled and non-scaled data
build_pll_lm_long <- function(lm_res_file, lm_death_res_file,
                              lm_res_phys_file, lm_death_res_phys_file) {
  pll_lm_pheno_res <- readRDS(file.path(local_filepath, lm_res_phys_file))
  pll_lm_death_pheno_res <- readRDS(file.path(local_filepath, lm_death_res_phys_file))

  pll_lm_aging <- readRDS(file.path(local_filepath, lm_res_file)) %>%
    dplyr::select(-method, compoundName) %>%
    dplyr::bind_rows(pll_lm_pheno_res)

  pll_lm_dying <- readRDS(file.path(local_filepath, lm_death_res_file)) %>%
    dplyr::select(-method, compoundName) %>%
    dplyr::bind_rows(pll_lm_death_pheno_res)

  dplyr::bind_rows(
    pll_lm_aging %>%
      dplyr::mutate(
        model_term = ifelse(model_term == "PLL", "PLL_Aging", as.character(model_term)),
        model = gsub("PLL", "PLL_Aging", model),
        phase = "Aging"
      ),
    pll_lm_dying %>%
      dplyr::mutate(
        model_term = ifelse(model_term == "PLL", "PLL_Dying", as.character(model_term)),
        model = gsub("PLL", "PLL_Dying", model),
        phase = "Dying"
      )
  ) %>%
    dplyr::left_join(., name_conversion_use, by = "feature_id") %>%
    dplyr::left_join(., pll_gam_cluster_identities, by = c("feature_id" = "trait")) %>%
    dplyr::rows_update(
      physio_key %>% dplyr::select(feature_id, name_use),
      by = "feature_id"
    ) %>%
    dplyr::filter(!is.na(name_use)) %>%
    dplyr::mutate(
      model = make_names_pretty(model),
      model_term = make_names_pretty(model_term)
    ) %>%
    dplyr::select(-padj) %>%
    fdr_multi(
      pval_var = "p_value_coef",
      nest_vars = c("phase", "model_term", "modality")
    )
}

# Build each and format as table_s6
make_table_s4 <- function(pll_lm_long, normalization) {
  pll_lm_long %>%
    dplyr::mutate(
      Platform = stringr::str_to_title(modality),
      Platform = factor(Platform, levels = c("Metabolomics", "Lipidomics", "Proteomics", "Physiological")),
      Normalization = normalization
    ) %>%
    dplyr::select(Normalization, Platform,
      Compound = name_use, Model = model,
      Phase = phase, Term = model_term, Estimate = coef,
      Est_SE = se_coef, Pvalue = p_value_coef, FDR = padj,
      Cluster = cluster
    )
}

table_s4 <- dplyr::bind_rows(
  make_table_s4(
    build_pll_lm_long(
      "20260220-Linear-Scaled-MM-PLL-Aging-Results.Rds",
      "20260220-Linear-Scaled-MM-PLL-Dying-Results.Rds",
      "20250528-Linear-MM-PLL-Aging-Phenotype-Results.Rds",
      "20250528-Linear-MM-PLL-Dying-Phenotype-Results.Rds"
    ),
    "Z-score"
  ),
  make_table_s4(
    build_pll_lm_long(
      "20250424-Linear-MM-PLL-Aging-Results.Rds",
      "20250424-Linear-MM-PLL-Dying-Results.Rds",
      "20250424-Linear-MM-PLL-Aging-Phenotype-Results.Rds",
      "20250424-Linear-MM-PLL-Dying-Phenotype-Results.Rds"
    ),
    "Log2 Relative"
  )
)

# write.csv(table_s4,
#           file.path(local_filepath, "supplemental_tables_final", "Table_S4_PLL_Aging_Dying_Model.csv"),
#           row.names = FALSE)
```

# Figure 3

## A: Clusters

1. Calculate weights for each data point, inversely proportional to the width of the confidence interval.
2. Group the data by cluster and PLL. Within each group, I compute the weighted mean of `scaled_fit`, `scaled_lower_ci`, and `scaled_upper_ci`, using the previously calculated weights.

Clusters have been manually re-identified so that the the plot shows the clusters next to each other with opposing directionality.

```{r, fig.width = 3.8, fig.height = 1.8}
# Scaling the data
scaled_data <- pll_gam_fits %>%
  dplyr::filter(significant) %>%
  dplyr::group_by(trait) %>%
  dplyr::group_modify(~ docr_scale_trajectory(.x)) %>%
  dplyr::ungroup()

# Compute average trajectory
average_trajectories <- scaled_data %>%
  dplyr::group_by(cluster, PLL) %>%
  dplyr::summarize(
    average_scaled_fit = mean(scaled_fit, na.rm = TRUE),
    average_scaled_lower_ci = mean(scaled_lower_ci, na.rm = TRUE),
    average_scaled_upper_ci = mean(scaled_upper_ci, na.rm = TRUE),
    .groups = "drop"
  )

count_data <- pll_gam_summaries %>%
  dplyr::filter(significant) %>%
  dplyr::distinct(cluster, name_use) %>%
  dplyr::group_by(cluster) %>%
  dplyr::summarise(count = n(), .groups = "drop") %>%
  dplyr::mutate(label_print = paste0("n = ", count))

# Visualization
g <- ggplot2::ggplot(
  average_trajectories,
  ggplot2::aes(
    x = PLL,
    y = average_scaled_fit,
    color = factor(cluster)
  )
) +
  ggplot2::geom_line() +
  ggplot2::geom_text(
    data = count_data,
    mapping = aes(x = Inf, y = Inf, label = label_print),
    color = "black",
    hjust = 1.2,
    vjust = 1.55,
    size = 2.2
  ) +
  ggplot2::geom_ribbon(
    ggplot2::aes(
      ymin = average_scaled_lower_ci,
      ymax = average_scaled_upper_ci,
      fill = factor(cluster)
    ),
    alpha = 0.2,
    color = NA
  ) +
  ggplot2::labs(
    title = "Lifespan trajectory cluster average",
    x = "PLL",
    y = "Scaled Predicted Fit",
    color = "Cluster",
    fill = "Cluster"
  ) +
  ggplot2::facet_wrap(~ factor(cluster), ncol = 5) +
  docr_ggplot_theme(
    x_text_angle = 0,
    x_text_hjust = 0.5
  ) +
  theme(
    strip.text = element_text(margin = margin(0, 2, 2, 2)),
    panel.spacing.y = unit(3, "pt")
  )
print(g)

# docr_ggsave(g, "Fig_3A_ClusterTrajectories", 1.9, 3.9, local_filepath = local_filepath)
```

## B: PMC

```{r, fig.width = 1.35, fig.height = 1.5}
# filter the data and calculate the median
filtered_data <- pll_gam_summaries %>%
  dplyr::filter(!is.na(cluster), extrema_trait != 1) %>%
  dplyr::distinct(cluster, name_use, extrema_trait)

median_val <- median(filtered_data$extrema_trait, na.rm = TRUE)

# plot
g <- ggplot(filtered_data, aes(x = extrema_trait)) +
  geom_histogram(
    color = "black",
    fill = "white",
    binwidth = 0.01
  ) +
  geom_vline(
    xintercept = median_val,
    color = "red",
    linewidth = 1
  ) +
  annotate(
    geom = "label",
    x = 0.6,
    y = 74,
    label = paste("Median =", round(median_val, 3)),
    color = "black",
    size = 5 / ggplot2::.pt
  ) +
  docr_ggplot_theme(x_text_angle = 0) +
  labs(
    title = paste0("Maximum Curvature"),
    x = "PMC",
    y = ""
  ) +
  ggplot2::annotate("text",
    x = 0.40, y = 125,
    label = paste0("n = ", nrow(filtered_data)),
    size = 5 / ggplot2::.pt
  ) +
  ggplot2::annotate("text",
    x = 0.53, y = 112,
    label = "non-linear slopes",
    size = 5 / ggplot2::.pt
  )

g

# docr_ggsave(g, "Fig_3C_MaxInflection_short", 1.5, plot_width = 1.35, local_filepath = local_filepath)
```

## C: Real Data

```{r, fig.width = 1.2, fig.height = 1.5}
death_shift <- TRUE

top_traits <- c("sp__P13020__GELS")

for (trait_use in top_traits) {
  one_aging <- pll_lm_aging %>%
    dplyr::filter(feature_id == trait_use)

  cluster_use <- pll_gam_summaries %>%
    dplyr::filter(trait == trait_use) %>%
    dplyr::pull(cluster)
  cluster_color <- scales::hue_pal()(10)[as.numeric(cluster_use)]

  one_aging_p <- one_aging %>%
    dplyr::filter(model_term == "PLL") %>%
    dplyr::pull(padj)

  one_aging <- one_aging %>%
    dplyr::select(model_term, coef) %>%
    tibble::deframe()

  one_dying <- pll_lm_dying %>%
    dplyr::filter(feature_id == trait_use) %>%
    dplyr::select(model_term, coef) %>%
    tibble::deframe()

  new_data <- data_use %>%
    dplyr::filter(
      trait_id == trait_use,
      age_years != "year1"
    ) %>%
    tidyr::drop_na(trait_value)

  custom_title <- new_data$name_use[1]

  if (custom_title == "C9") {
    custom_ggplot_title <- "Complement comp. 9"
  } else if (custom_title == "GELS") {
    custom_ggplot_title <- "Gelsolin"
  } else if (custom_title == "ITIH3") {
    custom_ggplot_title <- "ITI-HC3"
  } else {
    custom_ggplot_title <- custom_title
  }

  # set hex stuff
  if (custom_title == "2-Oxoglutarate") {
    custom_bins <- c(0.025, diff(range(new_data$trait_value, na.rm = TRUE)) / 40)
  } else if (custom_title == "N-acetylaspartate") {
    custom_bins <- c(0.025, diff(range(new_data$trait_value, na.rm = TRUE)) / 25)
  } else {
    custom_bins <- c(0.045, diff(range(new_data$trait_value, na.rm = TRUE)) / 20)
  }

  line_data <- tibble::tibble(
    PLL_aging = seq(min(new_data$PLL, na.rm = TRUE), 0.845, length.out = 10),
    PLL_dying = seq(0.85, max(new_data$PLL, na.rm = TRUE), length.out = 10)
  )

  # median bodyweight is 32 grams
  # diet 2D captures the median diet effect
  line_data <- line_data %>%
    dplyr::mutate(
      aging_pred =
        one_aging[["(Intercept)"]] +
          one_aging[["BW_Loess"]] * 32 +
          one_aging[["PLL"]] * PLL_aging,
      dying_pred =
        one_dying[["(Intercept)"]] +
          one_dying[["BW_Loess"]] * 32 +
          one_dying[["PLL"]] * PLL_dying
    )

  if (death_shift) {
    aging_endpoint <- line_data %>%
      dplyr::filter(PLL_aging == max(PLL_aging)) %>%
      dplyr::pull(aging_pred)
    dying_startpoint <- line_data %>%
      dplyr::filter(PLL_dying == min(PLL_dying)) %>%
      dplyr::pull(dying_pred)
    shift <- aging_endpoint - dying_startpoint
    line_data <- line_data %>%
      dplyr::mutate(dying_pred = dying_pred + shift)
  }

  gam_fit <- pll_gam_fits %>%
    dplyr::filter(trait %in% trait_use)

  g <- ggplot2::ggplot(new_data) +
    docr_ggplot_theme(
      x_text_angle = 0,
      x_text_hjust = 0.5
    ) +
    geom_hex(aes(x = PLL, y = trait_value), binwidth = custom_bins) +
    labs(
      title = paste0(custom_ggplot_title, ", Cluster ", cluster_use),
      y = "Log2 Abundance"
    ) +
    scale_fill_gradientn(colors = c("white", cluster_color)) +
    geom_line(
      data = gam_fit,
      mapping = aes(
        x = PLL,
        y = fit
      ),
      color = cluster_color,
      size = 1
    ) +
    geom_line(
      data = line_data, aes(x = PLL_aging, y = aging_pred),
      color = "black", size = 0.7, alpha = 0.6
    ) +
    geom_line(
      data = line_data, aes(x = PLL_dying, y = dying_pred),
      color = "black", size = 0.7, alpha = 0.6
    ) +
    docr_ggplot_line(x = 0.85)

  print(g)

  # docr_ggsave(g, "Fig_3C_ExAgingDying", 1.5, 1.2, local_filepath = local_filepath)
}
```

## D: Aging vs. Dying Scatter

```{r, fig.width = 1.5, fig.height = 1.5}
# PLL Wide
gplot <- pll_lm_long %>%
  dplyr::filter(model_term %in% c("PLL_Aging", "PLL_Dying")) %>%
  dplyr::select(model_term, modality, cluster, name_use, padj, coef) %>%
  tidyr::pivot_wider(
    names_from = model_term,
    values_from = c("padj", "coef")
  )

p_thresh <- 0.01
sd_dying <- sd(gplot$coef_PLL_Dying, na.rm = TRUE)
sd_aging <- sd(gplot$coef_PLL_Aging, na.rm = TRUE)

gplot <- gplot %>%
  dplyr::filter(!is.na(cluster)) %>%
  dplyr::mutate(
    color_use = ifelse((padj_PLL_Aging < p_thresh | padj_PLL_Dying < p_thresh), cluster, 11),
    color_use = as.factor(color_use)
  )

custom_colors <- c(scales::hue_pal()(10), "grey80")
names(custom_colors) <- as.character(1:11)

g1 <- ggplot2::ggplot(
  gplot,
  aes(
    y = coef_PLL_Aging,
    x = coef_PLL_Dying,
    color = color_use
  )
) +
  ggplot2::geom_point(
    alpha = 0.5,
    size = 0.6
  ) +
  docr_ggplot_line(x = 0) +
  docr_ggplot_line(y = 0) +
  scale_color_manual(values = custom_colors) +
  ggplot2::labs(
    title = paste0("Slopes, FDR < ", p_thresh),
    x = "Dying, Post-0.85 PLL",
    y = "Aging, Pre-0.85 PLL"
  ) +
  docr_ggplot_theme(
    x_text_angle = 0,
    x_text_hjust = 0.5
  )
print(g1)

# docr_ggsave(g1, paste0("Fig_3D_Aging_Dying_Coefs_small"), 1.5, 1.5, local_filepath = local_filepath)
```

### Text: Calculations

```{r}
p_thresh <- 0.01
calcs <- gplot %>%
  dplyr::mutate(sig_group = dplyr::case_when(
    padj_PLL_Aging < p_thresh & padj_PLL_Dying < p_thresh & sign(coef_PLL_Aging) == sign(coef_PLL_Dying) ~ "Both Significant, Correlated",
    padj_PLL_Aging < p_thresh & padj_PLL_Dying < p_thresh & sign(coef_PLL_Aging) != sign(coef_PLL_Dying) ~ "Both Significant, Anti-Correlated",
    padj_PLL_Aging < p_thresh & padj_PLL_Dying >= p_thresh ~ "Significant in Aging Only",
    padj_PLL_Aging >= p_thresh & padj_PLL_Dying < p_thresh ~ "Significant in Dying Only",
    TRUE ~ "sig_in_none"
  ))

table(calcs$sig_group)
```

## E, F, H: Heatmaps

```{r}
cluster_annotation <- pll_lm_long %>%
  dplyr::filter(!is.na(cluster)) %>%
  dplyr::distinct(name_use, Cluster = cluster) %>%
  tibble::column_to_rownames("name_use")

cluster_colors <- scales::hue_pal()(10)
names(cluster_colors) <- as.character(1:10)

n_thresh <- 40

physio_terms <- c(
  "Hematocrit %", "Hemoglobin conc. CV", "Hemoglobin",
  "Neutrophil-lymphocyte ratio", "RBC count", "% Lymphocytes",
  "% Neutrophils", "RBC volume CV", "Lymphocytes (% viable)"
)

top_terms <- function(term) {
  pll_lm_long %>%
    dplyr::filter(
      model_term == term,
      !grepl("TG", name_use),
      !is.na(cluster),
      modality != "physiological",
      !is.na(padj)
    ) %>%
    dplyr::arrange(padj) %>%
    dplyr::slice_head(n = n_thresh) %>%
    dplyr::pull(name_use) %>%
    unique()
}

list_loop <- list(
  `Top Aging` = top_terms("PLL_Aging"),
  `Top Dying` = top_terms("PLL_Dying"),
  `Top Physio` = physio_terms
)

# Compounds to bold:
bold_labels <- intersect(list_loop$`Top Aging`, list_loop$`Top Dying`)

annotation_col <- docr_pheatmap_annotations(
  pll_lm_long,
  `PLL Slope` = c("Aging" = "Aging", "Dying" = "Dying")
)

annotation_colors <- docr_pheatmap_colors(
  annotation_col
) %>%
  c(., list(Cluster = cluster_colors))

for (ll in names(list_loop)) {
  temp_terms <- list_loop[[ll]]
  temp_plot_use <- pll_lm_long %>%
    dplyr::filter(model_term %in% c("PLL_Aging", "PLL_Dying")) %>%
    dplyr::filter(name_use %in% temp_terms) %>%
    dplyr::mutate(padj = ifelse(is.na(padj), 1, padj))

  ph <- docr_pheatmap(
    temp_plot_use,
    main = ll,
    y_axis_var = "name_use",
    annotation_row = cluster_annotation,
    cluster_rows = TRUE,
    show_colnames = TRUE,
    show_rownames = TRUE,
    fontsize = 5,
    fontsize_row = 5,
    fontsize_number = 9,
    sig_symbol = "\u00B7",
    sig_numbers = c(0.001, 0.01, 0.05),
    annotation_colors = annotation_colors,
    display_var = "padj",
    pvalue_stars = TRUE,
    cellwidth = 11,
    cellheight = 4.5,
    treeheight_row = 0,
    angle_col = 45
  )

  # bold overlapping traits
  gt <- ph$gtable
  row_names_grob <- gt$grobs[[which(gt$layout$name == "row_names")]]
  labels <- row_names_grob$label
  bold_idx <- which(labels %in% bold_labels)
  fontfaces <- rep("plain", length(labels))
  fontfaces[bold_idx] <- "bold"
  row_names_grob$gp <- gpar(fontface = fontfaces)
  gt$grobs[[which(gt$layout$name == "row_names")]] <- row_names_grob
  grid.newpage()
  grid.draw(gt)

  # docr_ggsave(plot_object = gt,
  #             plot_file_name = paste0("Fig_3_Heatmap_P_B_", ll),
  #             plot_height = nrow(temp_plot_use) * 0.015 + 5,
  #             plot_width = 4,
  #             local_filepath = local_filepath)
}
```

# GO Enrichment

### Metabolites: INCHII enrichment

```{r}
enrich_data_metab <- pll_lm_long %>%
  dplyr::filter(
    modality %in% c("metabolomics", "lipidomics"),
    model_term %in% c("PLL_Aging", "PLL_Dying")
  ) %>%
  dplyr::mutate(sign_coef = sign(coef)) %>%
  dplyr::distinct(feature_id, model_term, sign_coef, padj) %>%
  dplyr::inner_join(inchii_key, by = "feature_id")

inchi_universe <- enrich_data_metab %>%
  dplyr::distinct(inchi_key) %>%
  dplyr::pull() %>%
  {
    gsub("\\xca", "E", ., useBytes = TRUE)
  } # fix symbol

enrich_metab <- data.frame()

for (tm in unique(enrich_data_metab$model_term)) {
  for (sc in unique(enrich_data_metab$sign_coef)) {
    enrich_data_temp <- enrich_data_metab %>%
      dplyr::filter(
        model_term == tm,
        sign_coef == sc,
        padj < 0.01
      ) %>%
      dplyr::pull(inchi_key) %>%
      unique()

    res_temp <- docr_inchikey_enrichment(
      input_inchi_vector = enrich_data_temp,
      input_universe = inchi_universe,
      go_names_filepath = go_names_filepath,
      go_inchi_key_filepath = go_inchi_key_filepath
    ) %>%
      dplyr::mutate(
        sign_coef = sc,
        model_term = tm
      ) %>%
      dplyr::ungroup() %>%
      # FDR calculation to match proteins as used in clusterProfiler::enricher()
      dplyr::mutate(qvalue = qvalue::qvalue(Pvalue,
        lambda = 0.05,
        pi0.method = "bootstrap"
      )$qvalues)
    enrich_metab <- enrich_metab %>%
      dplyr::bind_rows(res_temp)
  }
}
```

### Proteins: clusterProfiler

```{r}
# Get entrezIDs for all proteins
go_data <- pll_lm_long %>%
  dplyr::filter(
    modality == "proteomics",
    model_term %in% c("PLL_Aging", "PLL_Dying")
  ) %>%
  dplyr::mutate(sign_coef = sign(coef)) %>%
  dplyr::distinct(name_use, gene_name, model_term, sign_coef, padj) %>%
  dplyr::mutate(entrez_id = suppressMessages(
    AnnotationDbi::mapIds(
      org.Mm.eg.db,
      keys = .$gene_name,
      column = "ENTREZID",
      keytype = "SYMBOL",
      multiVals = "first"
    )
  )) %>%
  dplyr::filter(!is.na(entrez_id))

# Define all proteins detected in this dataset
my_entrez_universe <- unique(go_data$entrez_id)

# Make full TERM2GENE dataframe object for clustering
msigdbr_TERM2GENE <- msigdbr::msigdbr(
  species = "mouse",
  db_species = "MM",
  collection = "M5"
) %>%
  dplyr::select(gs_name, ncbi_gene)

# Initialized dataframe
enrich_proteins <- data.frame()

# RUN GO ENRICHMENT, for
# (1) each sign (up or down regulated)
# (2) each term, aging or dying
for (i in unique(go_data$sign_coef)) {
  for (j in unique(go_data$model_term)) {
    # Extract significant entrez IDs for group
    entrez_ids_sig <- go_data %>%
      dplyr::filter(
        model_term == j,
        sign_coef == i,
        padj < 0.01
      ) %>%
      dplyr::arrange(desc(abs(sign_coef))) %>%
      dplyr::pull(entrez_id)

    # Run cluster enrichment
    test <- clusterProfiler::enricher(
      gene = entrez_ids_sig,
      universe = my_entrez_universe,
      TERM2GENE = msigdbr_TERM2GENE
    )

    # Clean results object
    export_df <- test@result %>%
      dplyr::mutate(
        sign_coef = i,
        model_term = j,
        Overlap = Count,
        All = length(my_entrez_universe),
        Input = length(entrez_ids_sig),
        GO_term = Description,
        Process = as.integer(str_split(BgRatio, pattern = "/", simplify = TRUE)[, 1])
      ) %>%
      dplyr::select(GO_term, Process, Overlap, All, Input,
        `Fold Enrichment` = FoldEnrichment,
        Pvalue = pvalue,
        qvalue,
        sign_coef,
        model_term,
        geneID
      )

    # Bind to all cluster data
    enrich_proteins <- enrich_proteins %>%
      dplyr::bind_rows(export_df)
  }
}
```

### Plot Pathways
 
#### Jaccard de-duplication function

```{r}
# Operates on a df with columns: GO_term, sign_coef, Overlap_Compounds, FE_use
# Splits compound string by `sep`, filters greedily within each sign_coef group
# keeping the term with highest |FE| when two terms exceed jaccard_threshold
jaccard_dedup <- function(df, compound_col, sep, threshold = 0.5) {
  jaccard <- function(a, b) length(intersect(a, b)) / length(union(a, b))

  df %>%
    dplyr::group_by(sign_coef) %>%
    # Within each sign group, sort by |FE| descending so we keep strongest terms
    dplyr::arrange(dplyr::desc(abs(FE_use)), .by_group = TRUE) %>%
    dplyr::group_modify(~ {
      terms <- .x$GO_term
      compounds <- purrr::map(
        .x[[compound_col]],
        ~ stringr::str_split(.x, sep)[[1]]
      )
      names(compounds) <- terms

      kept <- c()
      for (term in terms) {
        if (length(kept) == 0) {
          kept <- c(kept, term)
          next
        }
        sims <- purrr::map_dbl(kept, ~ jaccard(compounds[[term]], compounds[[.x]]))
        if (all(sims < threshold)) kept <- c(kept, term)
      }
      dplyr::filter(.x, GO_term %in% kept)
    }) %>%
    dplyr::ungroup() %>%
    dplyr::distinct(GO_term, sign_coef)
}
```

## G:  Bar Plot 

```{r, fig.width = 3, fig.height = 5.5}
jaccard_threshold <- 0.5
qvalue_threshold <- 0.1

# make colors
bar_colors <- c(
  "Enriched Aging"  = "#FF918A",
  "Enriched Dying"  = "#D63B24",
  "Depleted Aging"  = "#7E88BA",
  "Depleted Dying"  = "#21336B"
)

# clean proteins
enrich_proteins_clean <- enrich_proteins %>%
  dplyr::filter(grepl("^GOBP_", GO_term)) %>%
  dplyr::mutate(
    omics = "P",
    Overlap_Compounds = purrr::map_chr(
      stringr::str_split(geneID, "/"),
      ~ stringr::str_c(sort(.x), collapse = "/")
    )
  )

# clean metabolites and add proteins
enrich_all_clean <- enrich_metab %>%
  dplyr::mutate(
    omics = "M",
    Overlap_Compounds = purrr::map_chr(
      stringr::str_split(Overlap_Compounds, "\\|"),
      ~ stringr::str_c(sort(.x), collapse = "/")
    )
  ) %>%
  dplyr::bind_rows(enrich_proteins_clean) %>%
  dplyr::mutate(
    type = ifelse(grepl("Aging", model_term), "Aging", "Dying"),
    FE_use = `Fold Enrichment` * sign_coef
  )

# get go terms to plot
go_term_use_all <- enrich_all_clean %>%
  dplyr::filter(qvalue < qvalue_threshold, Overlap > 3) %>%
  dplyr::group_by(Overlap_Compounds) %>%
  dplyr::arrange(Process) %>%
  dplyr::slice_head(n = 1) %>%
  dplyr::ungroup() %>%
  jaccard_dedup(compound_col = "Overlap_Compounds", sep = "/", threshold = jaccard_threshold)

combined <- enrich_all_clean %>%
  dplyr::inner_join(go_term_use_all, by = c("GO_term", "sign_coef")) %>%
  dplyr::mutate(
    GO_term_label = dplyr::case_when(
      omics == "M" ~ stringr::str_replace(GO_name, "^(.)(?![[:upper:]])", stringr::str_to_upper),
      omics == "P" ~ GO_term %>% gsub("^GOBP_", "", .) %>% gsub("_", " ", .) %>% stringr::str_to_sentence(),
      TRUE ~ GO_term
    )
  ) %>%
  dplyr::filter(qvalue < qvalue_threshold) %>%
  dplyr::mutate(
    y_label = paste0(GO_term_label, " (", omics, ")"),
    group = dplyr::case_when(
      sign_coef == 1 & type == "Aging" ~ "Enriched Aging",
      sign_coef == 1 & type == "Dying" ~ "Enriched Dying",
      sign_coef == -1 & type == "Aging" ~ "Depleted Aging",
      sign_coef == -1 & type == "Dying" ~ "Depleted Dying"
    )
  )

# make facets
sig_facet <- combined %>%
  dplyr::distinct(y_label, sign_coef, type) %>%
  dplyr::group_by(y_label, sign_coef) %>%
  dplyr::summarise(
    has_aging = any(type == "Aging"),
    has_dying = any(type == "Dying"),
    .groups   = "drop"
  ) %>%
  dplyr::mutate(
    sig_group = dplyr::case_when(
      has_aging & has_dying ~ "Both",
      has_dying ~ "Dying only",
      has_aging ~ "Aging only"
    ),
    sig_group = factor(sig_group, levels = c("Both", "Aging only", "Dying only"))
  ) %>%
  dplyr::select(y_label, sign_coef, sig_group)

combined_plot <- combined %>%
  dplyr::left_join(sig_facet, by = c("y_label", "sign_coef"))

# make plot
label_overrides <- c(
  "Humoral immune response mediated by circulating immunoglobulin (P)" = "Humoral immune response mediated by Ig (P)",
  "tRNA aminoacylation for mitochondrial protein translation (M)" = "Mitochondrial tRNA aminoacylation (M)",
  "Negative regulation of protein containing complex assembly (P)" = "Neg. regulation of protein complex assembly (P)",
  "Proteasome mediated ubiquitin dependent protein catabolic process (P)" = "UPS-mediated protein degradation (P)",
  "Nucleobase-containing small molecule biosynthetic process (M)" = "Nucleobase-containing biosynthesis (M)",
  "Positive regulation of protein containing complex assembly (P)" = "Pos. regulation of protein complex assembly (P)",
  "Negative regulation of intracellular signal transduction (P)" = "Neg. regulation of intracellular signal transduction (P)"
)

plot_data <- combined_plot %>%
  dplyr::mutate(
    group = factor(group, levels = names(bar_colors))
  ) %>%
  dplyr::mutate(y_label = dplyr::recode(y_label, !!!label_overrides))

y_order <- plot_data %>%
  dplyr::group_by(y_label, sig_group, sign_coef) %>%
  dplyr::summarise(
    dying_fe = suppressWarnings(max(abs(FE_use[type == "Dying"]), na.rm = TRUE)),
    aging_fe = suppressWarnings(max(abs(FE_use[type == "Aging"]), na.rm = TRUE)),
    .groups  = "drop"
  ) %>%
  dplyr::mutate(
    dying_fe = ifelse(is.infinite(dying_fe), 0, dying_fe),
    aging_fe = ifelse(is.infinite(aging_fe), 0, aging_fe)
  ) %>%
  dplyr::arrange(sig_group, desc(sign_coef), dying_fe, aging_fe) %>%
  dplyr::pull(y_label) %>%
  unique()

plot_data <- plot_data %>%
  dplyr::mutate(y_label = factor(y_label, levels = y_order))

g_combined <- ggplot(plot_data) +
  docr_ggplot_theme(
    x_text_angle    = 0,
    x_text_hjust    = 0.5,
    legend_position = "none"
  ) +
  geom_col(
    data = ~ dplyr::filter(., sig_group == "Both"),
    aes(y = y_label, x = FE_use, fill = group, color = group),
    position = position_dodge2(preserve = "single", padding = 0.1),
    width = 0.7,
    linewidth = 0.3
  ) +
  geom_col(
    data = ~ dplyr::filter(., sig_group != "Both"),
    aes(y = y_label, x = FE_use, fill = group, color = group),
    position = position_dodge2(preserve = "single", padding = 0.1),
    width = 0.7,
    linewidth = 0.3
  ) +
  scale_fill_manual(name = "", values = bar_colors, drop = FALSE) +
  scale_color_manual(name = "", values = bar_colors, drop = FALSE) +
  scale_y_discrete(position = "right") +
  docr_ggplot_line(x = 0) +
  facet_grid(
    rows   = vars(sig_group),
    scales = "free_y",
    space  = "free_y",
    switch = "y"
  ) +
  labs(title = "GO Enrichment", y = "", x = expression("Fold Enrichment")) +
  theme(
    axis.text.y = element_text(size = 5),
    axis.text.y.right = element_text(size = 5),
    strip.text.y.left = element_text(
      size = 6, hjust = 0.5, angle = 90,
      margin = margin(2, 2, 2, 2)
    ),
    plot.title = element_text(margin = margin(b = 2)),
    panel.spacing = unit(2, "pt")
  ) +
  ggplot2::guides(
    fill  = guide_legend(nrow = 2, byrow = TRUE),
    color = guide_legend(nrow = 2, byrow = TRUE)
  )

print(g_combined)

# docr_ggsave(g_combined, "Fig_3G_GO_All", plot_height = 5.3, 3, local_filepath = local_filepath)
```

# Supplement

## Table S5

```{r}
conv_key <- inchii_key %>%
  dplyr::mutate(
    inchi_key = gsub("\\xca", "E", inchi_key, useBytes = TRUE),
    inchi_key = sub("-.*", "", inchi_key)
  ) %>%
  dplyr::left_join(name_conversion_use, by = "feature_id") %>%
  dplyr::select(name_use, go_key = inchi_key) %>%
  dplyr::bind_rows(go_data %>% dplyr::distinct(name_use, go_key = entrez_id))

table_s7 <- plot_data %>%
  dplyr::mutate(
    Molecules = sapply(strsplit(Overlap_Compounds, "/"), function(keys) {
      paste(conv_key$name_use[match(keys, conv_key$go_key)], collapse = ", ")
    })
  ) %>%
  dplyr::select(GO_term,
    GO_name = y_label, Overlap, Process, Input, All,
    Fold_Enrichment = `Fold Enrichment`, FE_use, Group = group,
    Facet = sig_group, Pvalue, FDR = qvalue, Molecules
  ) %>%
  dplyr::arrange(Facet, FE_use)

# write.csv(table_s5,
#           file.path(local_filepath, "supplemental_tables_final", "Table_S5_GO_Enrichment.csv"),
#           row.names = FALSE)
```

## 2B: Cluster Data

```{r, fig.width = 5.9, fig.height = 2.3}
# Potential to overlap biological information on this
count_data <- pll_gam_summaries %>%
  dplyr::filter(significant) %>%
  dplyr::group_by(cluster) %>%
  dplyr::summarise(count = n(), .groups = "drop") %>%
  dplyr::mutate(
    label_print = paste0("n = ", count),
    cluster = as.factor(cluster)
  )

g <- ggplot2::ggplot(pll_gam_summaries %>% dplyr::filter(significant)) +
  ggplot2::geom_bar(aes(
    y = stringr::str_to_title(modality),
    fill = cluster
  )) +
  ggplot2::labs(
    title = "Number of features per cluster by modality",
    x = "Number of Features",
    y = ""
  ) +
  ggplot2::geom_text(
    data = count_data,
    mapping = aes(x = Inf, y = -Inf, label = label_print),
    color = "black",
    hjust = 1.2,
    vjust = -1,
    size = 2.2
  ) +
  docr_ggplot_theme(
    x_text_angle = 0,
    x_text_hjust = 0.5
  ) +
  ggplot2::facet_wrap(~cluster, ncol = 5)

print(g)
# docr_ggsave(plot_object = g,
#             "Fig_S4_ModalityN",
#             plot_width = 5.64,
#             plot_height = 2.3,
#             local_filepath = local_filepath)
```

